Action mechanism of a novel agrichemical quinofumelin against Fusarium graminearum

Version 1

Abstract

Abstract

Modern fungicides have made significant contributions to crop disease management, but the development of resistant fungal strains has caused their failure in disease control. Therefore, developing fungicides with novel action mechanisms is the most effective measure to manage resistance. Quinofumelin, a novel quinoline fungicide, exhibits exceptional antifungal activity against phytopathogens. However, there is currently no available information on its mechanism of action. Here, we used transcriptome and metabolome analysis to observe a co-enrichment pattern of differentially expressed genes (DEGs) and differentially accumulated metabolites (DAMs) within pyrimidine biosynthesis pathway (PBP), identifying down-regulation of dihydroorotate dehydrogenase (DHODH). Exogenous UMP, uridine or uracil (metabolites in PBP) successfully restored quinofumelin-induced inhibition of mycelial growth in Fusarium graminearum and Fusarium asiaticum. Additionally, the deletion of FgDHODHII was determined to be lethal; however, mycelial growth of ΔFgDHODHII mutants could be restored by adding UMP, uridine or uracil. These findings indicate that the deficiencies in FgDHODHII are functionally equivalent to complete inhibition of its activity by quinofumelin. Finally, molecular docking, surface plasmon resonance (SPR) and microscale thermophoresis (MST) results strongly support the precise interaction between quinofumelin and FgDHODHII. Collectively, these findings provide compelling evidence for the involvement of de novo uracil biosynthesis as mechanism of action for quinofumelin while identifying FgDHODHII as its specific target.

Introduction

Agrochemical fungicides have a history spanning over 3000 years in the control of plant diseases, and have undergone a development process from traditional protective fungicides with multiple action sites to modern selective fungicides with single action site1. The introduction of the first fully systemic seed treatment agent (carboxin, a succinate dehydrogenase inhibitor) in 1966 and the launch of the first broad-spectrum foliar systemic fungicide (benomyl, a tubulin polymerization inhibitor) in 1968 represent significant milestones in the advancement of modern selective fungicides with single action site2-3. Owing to their exceptional selectivity, potent activity, specific target site, and low toxicity, modern selective fungicides have gained widespread acceptance and witnessed an increasing market share year after year. However, like every coin has two sides, these types of fungicides encounter substantial challenges related to serious resistance issues when deployed in agricultural fields. Numerous reports have been published on plant pathogens developing resistance to modern selective fungicides, which often result in failures or reductions in the effectiveness of chemical control against plant diseases and lead to decreased crop yields and quality (https://www.frac.info/)4-6. Therefore, it is crucial to expedite the development of new-generation selective fungicides with innovative modes of action. Based on a single fungicide target, numerous modern selective fungicides can be developed. For instance, 24 commercialized available fungicides have been created so far based on the respiratory chain complex II as the target. Similarly, another 20 commercialized fungicides have been developed targeting the respiratory chain complex III. Additionally, there are already 37 commercialized fungicides that specifically target the fungicide target C14-demethylase in ergosterol biosynthesis (https://www.frac.info/). Therefore, discovering novel targets poses a significantly greater challenge compared to creating new fungicides within the existing categories.

According to the latest information released by the Fungicide Resistance Action Committee (FRAC), over 30 fungicide selective targets have been identified and acknowledged in the field of fungicide biological research spanning more than 130 years. Currently, these targets have been identified based on the successful development of fungicides and are all related proteins that are essential for pathogen growth or pathogenicity. Extensive identification and functional analysis of essential genes for growth or pathogenicity in plant pathogens have been achieved with rapid advancements in molecular biology techniques; however, there is no precedent for successfully developing commercialized fungicides targeting genes involved in growth or pathogenicity. The creation of highly active modern selective fungicides currently relies on the structural characteristics of existing targets and is accomplished through technologies such as AI drug discovery and design (AIDD), and computer aided drug design (CADD). The biological functions of drug targets are highly conserved, although variations may occur at critical amino acid sites during genetic evolution. Therefore, it is crucial to explore drug targets and analyze their structural characteristics in order to develop novel fungicides with enhanced activity and selectivity. Recently, the Fungicide Biology group from Nanjing Agricultural University proposed a new concept for targeting pesticides based on the principles of targeted medicine. The objective is to create specialized targeted pesticides that exhibit higher activity, greater selectivity, stronger specialization, lower toxicity, and environmental friendliness by leveraging the structural characteristics of pesticide targets. This approach aims to provide essential technical support for scientifically controlling crop diseases.

Quinofumelin (CAS 861647-84-9) is a novel quinoline fungicide developed by Mitsui Chemicals Co., Ltd. Japan. Due to their significant bioactivity, quinoline compounds and their derivatives have found extensive applications in medicine and pesticides7-9. Based on their biological and physiological characteristics, quinolines are classified as FRAC group 13 (https://www.frac.info/). However, the precise mode of action of quinolines against plant pathogens remains unclear. In addition, quinofumelin exhibits no cross-resistance to currently extensively used fungicides, indicating its unique action target against phytopathogenic fungi. Therefore, this study unveils that the mechanism of action for quinofumelin involves de novo synthesis of uracil that identifies dihydroorotate dehydrogenase as its specific target. This newly discovered selective target for antimicrobial agents provides a valuable resource for the design and development of targeted pesticides.

Results

Transcriptomic analysis of F. graminearum as affected by quinofumelin

To evaluate the quality of RNA-Seq data, we conducted a comprehensive analysis for quality control. In the principal component analysis, PC1 and PC2 accounted for 45.59% and 16.98% of the total variance, respectively. The distinct separation observed between the groups indicates significant differences in gene expression between the controls and experimental samples, which could be used for subsequent investigations (Supporting Information, Fig. S1). The clustering heat map effectively categorized the controls and experimental samples into distinct clusters based on their gene expression profiles, demonstrating robust biological replication with minimal intra-group variability and ensuring the reliability of the data. Moreover, notable inter-group dissimilarities highlighted variations among different treatment samples (Fig. 1c).

Analysis of transcriptome sequencing data in F. graminearum PH-1 treated with quinofumelin.

a Bar chart illustrating the comparison of differentially expressed genes between EC90 treatments and the control.

b Volcano plot depicting up-regulated DEGs as red scattered dots and down-regulated DEGs as blue scattered dots.

c Cluster heat maps displaying gene relative expression values using RNA-seq data scale-standardized values.

d Bubble map presenting the results of GO enrichment analysis for DEGs.

e Bubble map showcasing the findings of KEGG enrichment analysis for DEGs.

The differential analysis unveiled a total of 234 significantly differentially expressed genes (DEGs) between the control and experimental samples. Among these findings, there were identified as up-regulated for a count of 97 genes while down-regulation was observed in case of another set comprising around 137 genes (Fig. 1a). A graphical representation called “volcano plot” was employed to depict the statistical distribution pattern exhibited by these DEGs (Fig. 1b). To gain deeper insights into gene functionalities, Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) enrichment analysis were performed on these DEGs. The results showed that they are notably enriched in seventeen distinct GO terms spanning across two major categories; eleven molecular functions along with six processes (Fig. 1d). Furthermore, the enrichment analysis highlighted significant associations with various functional attributes including transition metal ion binding, oxidoreductase activity, modified amino acid binding, amide binding, monooxygenase activity, and heme binding. Moreover, the enriched biological processes primarily involved transcriptional regulation driven by nucleic-enabled mechanisms as well as monocarboxylic acid metabolism and RNA biosynthesis. Additionally, the KEGG enrichment analysis identified specific pathways associated with thiamine metabolism, tryptophan metabolism, nitrogen metabolism, amino acid sugar and nucleotide sugar metabolism, pantothenic acid and CoA biosynthesis, as well as nucleotide sugar production compound synthesis (Fig. 1e).

Metabolomic analysis of F. graminearum as affected by quinofumelin

In metabolomics principal component analysis, the contribution rate of PC1 was determined to be 76.6% in positive ion mode and 51.2% in negative ion mode, respectively. Moreover, the quality control (QC) samples exhibited highly concentrated results, indicating excellent instrument stability and remarkable data repeatability. Furthermore, a distinct separation between groups was observed, particularly prominent in the positive ion mode, indicating significant variations in metabolites among different groups that warrant further investigation (Supporting Information, Fig. S2a and S2b). The findings from Pearson correlation analysis were consistent with those obtained from principal component analysis (Supporting Information, Fig. S2c and S2d). Strong correlations were identified within samples belonging to the same group, while weak correlations were observed between different groups.

Metabolomics analysis revealed a total of 713 differential metabolites, with 14 up-regulated and 699 down-regulated in the positive ion mode, as well as 230 differential metabolites, including 103 up-regulated and 127 down-regulated in the negative ion mode (Supporting Information, Fig. 2a and Fig. S3a). Based on the classification of the identified 943 different metabolites, it was observed that amino acids and their derivatives accounted for the highest proportion (17%), followed by organic acids and their derivatives (11%), benzene and substituted derivatives (11%), aldehydes, ketones, esters (11%), and heterocyclic compounds (10%) (Fig. 2c). The volcano plot visually represents the distribution of statistical differences in metabolite abundance between the two groups (Supporting Information, Fig. 2b and Fig. S3b).

Difference analysis and enrichment analysis of metabolome data in F. graminearum PH-1 treated with quinofumelin.

a Bar chart illustrating the number of differential accumulation metabolites (DAMs) quantity in experimental groups (EC90) and control groups (CK) in positive ion mode.

b Volcanic map of all metabolites in positive ion mode. The red scatter points represent up-regulated DAMs, while the blue scatter points represent down-regulated DAMs.

c Classification and proportion of DAMs.

d Bubble diagram of differential accumulation metabolite KEGG enrichment pathway.

e Pyrimidine de novo biosynthesis pathway in Fusarium graminearum. The left box plots represent the relative content of 8 metabolite samples, while the right box represents the relative content of 6 gene samples. The color gradient from navy blue to firebrick red indicates a progression from low to high content.

KEGG pathways enriched by differential metabolites primarily include arachidonic acid metabolism, pyrimidine metabolism, alanine, aspartate and glutamate metabolism, purine metabolism, citrate cycle, arginine and proline metabolism, fructose and mannose metabolism, nitrogen metabolism, and metabolic pathways (Fig. 2d). The classification of enrichment pathways for differential metabolites revealed that the majority of these metabolites were associated with nucleotide metabolism (25%), amino acid metabolism (25%), carbohydrate metabolism (25%), and energy metabolism (12.5%) (Supporting Information, Fig. S3c).

Conjoint analysis of transcriptome and metabolome integration

Through transcriptome and metabolome data analysis, we observed a co-enrichment of differentially expressed genes and differentially accumulated metabolites in the pyrimidine metabolism pathway. In the expression calorigrams of metabolites and genes depicting the de novo biosynthesis pathway of pyrimidine, we identified down-regulation of dihydroorotate dehydrogenase (DHODH) in the experimental groups. This down-regulation hinders the conversion of dihydroorotate to orotate, thereby inhibiting the biosynthesis of downstream metabolites orotidine-5P and uracil, leading to their decreased abundance (Fig. 2e). Thus, we conclude that the action mechanism of quinofumelin may involve inhibition of the de novo biosynthesis pathway of pyrimidine.

Identification of DHODHII in F. graminearum

The FgDHODHII sequences in F. graminearum were identified by performing a BLAST search against the National Center for Biotechnology Information (NCBI) GenBank database, using Homo sapiens DHODHII amino acid sequences as a query. FgDHODHII is a gene consisting of 1419 bp without introns and is predicted to encode a protein consisting of 472 amino acids. Orthologs of FgDHODHII were obtained using the BLAST algorithm. The phylogenetic tree was constructed based on the amino acid sequences of DHODHII homologous protein with Mega X using the neighbor-joining method. Motif pattern information was generated using the MEME suite10, while functional domain information and protein coordinates were obtained through CD-search11. TBtools software was used for visualizing the tree representation12. All DHODHII proteins possess the dihydroorotate dehydrogenase domain (Fig. 3), indicating high conservation among fungi. Besides, these orthologs proteins would potentially interact with the fungicide in other pathosystems.

Phylogenic tree of DHODHII proteins.

The phylogenetic tree was constructed based on the amino acid sequences of DHODHII homologous protein with Mega X using the neighbor-joining method. The bootstrap values from 1000 replications are indicated on the branches. Motif pattern information was generated using the MEME suite12, while functional domain information and protein coordinates were obtained through CD-search. The final phylogenetic tree was visualized using TBtools. FgDHODHII was highlighted in red. The amino acid sequence of DHODHII from Alternaria alternata (AA0117_g542), Aspergillus nidulans (ANIA_05909), Aspergillus oryzae (AO1008_05965), Blumeria graminis (BLGH_00861), Botrytis cinerea (Bcin15g04150), Botrytis elliptica (BELL_0162g00150), Colletotrichum asianum (GQ607_007053), Colletotrichum fioriniae (CFIO01_03831), Colletotrichum higginsianum (CH063_10143), Corynespora cassiicola (BS50DRAFT_576390), Erysiphe necator (EV44_g2092), Erysiphe pulchra (EPUL_004094), Fusarium euwallaceae (BHE90_002825), Fusarium floridanum (CEP51_012005), Fusarium fujikuroi (FFMR_05040), Fusarium graminearum (FGSG_09678), Fusarium nygamai (FNYG_10990), Fusarium odoratissimum (FOIG_08270), Fusarium oxysporum (FOXG_08390), Fusarium pseudograminearum (FPSE_04283), Fusarium solani (NechaG15121), Fusarium verticillioides (FVEG_06288), Magnaporthe oryzae (MGG_08814), Neurospora crassa (NCU06532), Oidium neolycopersici (OnM2_089032), Pestalotiopsis fici (PFICI_04984), Podospora anserina (PODANS_6_7920), Puccinia graminis (PGTG_16739), Puccinia triticina (PTTG_02070), Rhizoctonia solani (RSOLAG22IIIB_08327), Schizosaccharomyces pombe (SPAC57A10.12c), Ustilaginoidea virens (UVI_02052880), Verticillium alfalfae (VDBG_01322), Verticillium dahliae (VDAG_07935), and Zymoseptoria tritici (Mycgr3G72548) were accessed in EnsemblFungi database.

Recovery test of mycelial growth suppressed by quinofumelin in Fusarium genus

The mycelial growth of F. graminearum PH-1 was significantly inhibited on CZA plates containing quinofumelin at concentrations of 1 μg/mL (Fig. 4). To investigate the potential recovery of mycelial growth by exogenous dihydroorotate, UMP, uridine, or uracil, CZA plates supplemented with quinofumelin were treated with these compounds. Despite the addition of exogenous dihydroorotate at 50 μg/mL failing to restore mycelial growth, the addition of exogenous UMP, uridine or uracil at 50 μg/mL successfully recovered quinofumelin-induced inhibition of mycelial growth. Similar results were observed in other strains of F. graminearum or F. asiaticum (Fig. 4 and Fig. S4). Thus, it can be concluded that quinofumelin effectively inhibits the pyrimidine biosynthesis pathways.

Recovery test of mycelial growth suppressed by quinofumelin.

All strains were incubated on CZA plates at 25°C for 3 days. F. graminearum strains PH-1, 512E-8, SX2117, XY-4, and R10. The left image shows the colony morphology, while the right image is a bar chart of colony diameters.

The absence of FgDHODHII is lethal for mycelial growth in F. graminearum

To further investigate the biological functions of FgDHODHII in F. graminearum, we employed a homologous recombination strategy to generate target gene deletion mutants. Through PCR analysis, two putative FgDHODHII deletion mutants were successfully identified. The ΔFgDHODHII mutants exhibited distinct phenotypic differences compared to the wild-type progenitor PH-1 and the respective complemented strain (Fig. 5a). These mutants displayed an inability to grow on CZA plate; however, their mycelial growth was restored when supplemented with 50 μg/mL of exogenous UMP, uridine or uracil, while no recovery was observed with 50 μg/mL of exogenous dihydroorotate (Fig. 5b).

Mycelial growth of FgDHODHII deletion mutants on CZA plates.

a Mycelial growth of the FgDHODHII deletion mutants and the parental strain was observed in the presence or absence of 50 μg/mL uridine.

b Recovery of mycelial growth of the FgDHODHII deletion mutants was observed in the presence of 50 μg/mL dihydroorotate, UMP, uridine or uracil. All strains were incubated at 25°C for 3 days.

Molecular docking

To facilitate the molecular docking process between FgDHODHII and quinofumelin, ten distinct binding models were generated using AutoDockTools-1.5.7 software, each accompanied by its respective binding energies calculation results. The most favorable conformational arrangement was determined based on its relative ranking in terms of the computed binding energy values and subsequently visualized through PyMOL 2.5.4 software (Fig. 6a). Notably, the obtained docking results revealed a robust interaction between FgDHODHII protein and quinofumelin as evidenced by a calculated negative free energy value of -6.8 kcal/mol.

Validation and analysis of quinofumelin”s binding affinity to FgDHODHII.

a Three-dimensional model of the wild-type FgDHODHII are depicted, with an enlarged view highlighting the binding site for quinofumelin on the corresponding FgDHODHII models.

b The SPR response (RU) analysis and fitting results demonstrate the interaction between quinofumelin and FgDHODHII.

c Microscale thermophoresis (MST) analysis to detect the binding of quinofumelin and FgDHODHII.

The precise interaction between FgDHODHII and quinofumelin was verified by Surface plasmon resonance analysis

Surface plasmon resonance (SPR) was utilized to evaluate the interaction affinity between quinofumelin and FgDHODHII in this study. The experiment was conducted using a BIAcore T200 (Cytiva) instrument for biomolecular interaction analysis. FgDHODHII was immobilized as the ligand on the chip surface, while quinofumelin continuously flowed through as the analyte solution. Dilute quinofumelin to different concentration gradients (62.5, 32.15, 15.63, 7.81, 3.91 μM), measure the binding energy, and then fit the binding equilibrium dissociation constant (Kd) value (Fig. 6b). The SPR results revealed that quinofumelin exhibited fast-binding and fast-dissociation modes when interacting with the target protein FgDHODHII. The calculated KD value was determined as 6.606×10−6 M, indicating specific binding of quinofumelin to FgDHODHII.

The interaction between FgDHODHII and quinofumelin was confirmed by microscale thermophoresis

In this study, microscale thermophoresis (MST) was further used to evaluate the interaction affinity between FgDHODHII and quinofumelin. The strength of the binding can be inferred from the Kd value. The experiment was conducted using MO.Control v1.6.1 and data analysis was performed using MO.Affinity Analysis v2.3. By introducing a fluorescent label GFP to the FgDHODHII protein while maintaining its concentration constant, non-fluorescent labeled small molecules were diluted into 16 concentration gradients. Subsequently, the protein solution and small molecule diluent were mixed at a ratio of 19:1 before being loaded into the capillary for further analysis in the machine. The MST assay demonstrated that FgDHODHII exhibited specific binding to quinofumelin in vitro, with a dissociation constant (Kd) of 0.471 ± 0.268 μM, whereas no binding was observed with phenamacril which lacks cross-resistance with quinofumelin (Fig. 6c).

Discussion

In over 130 years of research on fungicide biology, more than 30 novel targets for selective fungicides have been identified. Based on these newly discovered targets, over 200 modern selective fungicides have been developed. Consequently, the identification of novel targets for fungicides poses a significantly greater challenge compared to the development of new fungicides. Currently, modern selective fungicides with single action site are widely applied in agricultural practices for crop disease management. However, these specific types of fungicides encounter substantial obstacles associated with resistance issues when deployed in agricultural fields. Numerous studies have extensively documented cases of plant pathogens developing resistance to modern selective fungicides, resulting in failures or reductions in their efficacy against plant diseases and subsequently leading to decreased crop yields and quality4-6. For instance, Botrytis cinerea, a significant plant pathogenic fungus, has emerged as multi-resistant population that is resistant to all currently used modern selective fungicides. This poses a formidable challenge for the control of crop grey mold disease caused by B. cinerea. Therefore, there is an urgent need for the development of new fungicides with novel mechanisms of action to effectively manage multiple resistance.

The novel quinoline, quinofumelin, developed by Mitsui Chemicals Co., Ltd. Japan, has demonstrated remarkable antifungal activity against plant pathogens such as F. graminearum and Sclerotinia sclerotiorum7, 13. Moreover, our research group found that quinofumelin exhibits no cross-resistance to all types of currently used modern selective fungicides. These findings indicate that quinofumelin possesses a distinctive mode of action against plant pathogens. Based on their biological and physiological characteristics, quinolines are classified as FRAC group 13 (https://www.frac.info/). However, the precise mechanism by which quinolines act against plant pathogens remains unclear. In this study, we investigated the mechanism of action of quinofumelin on F. graminearum using a combined analysis of transcriptomics and metabolomics. The transcriptome data revealed that DEGs were involved in “tryptophan metabolism” and “amino sugar and nucleotide sugar metabolism”. Furthermore, the metabolome data identified DAMs enriched in “nucleotide metabolism” and “amino acid metabolism”. Interestingly, we observed a co-enrichment of DEGs and DAMs in the pyrimidine metabolism pathway. Specifically, we identified a down-regulation of dihydroorotate dehydrogenase (DHODH) in the experimental groups, which impeded the conversion of dihydroorotate to orotate. Consequently, this inhibition led to decreased biosynthesis of downstream metabolites orotidine-5P and uracil. Therefore, we conducted an experiment using F. graminearum standard strain PH-1 to quantify mycelial growth by supplementing exogenous uracil, which is the end product of the uracil biosynthesis pathway. This was performed in the presence of complete inhibition of mycelial growth caused by quinofumelin. Surprisingly, the addition of exogenous uracil restored the inhibited mycelial growth induced by quinofumelin (Fig. 4). Meanwhile, similar results were successfully verified with four other F. graminearum strains and five F. asiaticum strains (Fig. 4 and Fig. S4). Based on our depicted pyrimidine biosynthesis pathway (Fig. 2E), we also supplemented other intermediate products such as dihydroorotate, orotate, UMP, and uridine externally. Except for dihydroorotate and orotate, supplementation of exogenous UMP and uridine effectively restored mycelial growth inhibited by quinofumelin as well. These findings strongly support that quinofumelin acts through modulation of the uracil biosynthesis pathway.

As a key enzyme in the de novo pyrimidine biosynthesis pathway, DHODH has previously been investigated as a potential target for inhibiting infectious organisms and as an antitumor agent in mammals14-15. Inhibiting DHODH can effectively impede the synthesis of DNA, RNA, and glycoproteins, thereby regulating aberrant cell metabolism and proliferation to ultimately achieve therapeutic objectives16. Within mammalian cells, DHODH is closely associated with the mitochondrial membrane and plays a pivotal role in the respiratory chain through its interaction with ubiquinone17. Here, using amino acid sequences from Homo sapiens DHODHII as a reference query, we have identified FgDHODHII (FGSG_09678), a gene comprising of 1419 bp without any introns, which encodes 472 amino acids in F. graminearum. FgDHODHII possesses a typical DHOD-2 like domain (NCBI: cd04738) and belongs to class 2 DHODs. Class 2 DHODHs are membrane-bound enzymes, whereas class 1 enzymes are located in the cytosol18. The DHODH enzyme catalyzes a unique redox reaction in the de novo pyrimidine biosynthesis pathway19. The DHOD-2 like domain can bind orotate and flavin mononucleotide (FMN), facilitating the oxidation of dihydroorotate to orotate and the reduction of the FMN cofactor20. Based on the aforementioned findings, we have concluded that DHODHII functions as the target of action for quinofumelin against F. graminearum. Consequently, we performed a protoplast transformation assay to delete FgDHODHII in F. graminearum and further validate the precise target of action for quinofumelin. Typically, most fungicide targets are either lethal genes or essential pathogenic genes. Therefore, it is plausible that the inactivation of FgDHODHII may be lethal. In our genetic transformation experiment, exogenous uridine was supplemented to successfully obtain deletion mutants of the FgDHODHII gene. Fortunately, two ΔFgDHODHII mutants were successfully obtained. These mutants exhibited an inability to grow on PDA or CZA medium; however, mycelial growth was fully restored upon addition of uridine to the medium (Fig. 5A). Additionally, exogenous supplementation of UMP or uracil effectively restored mycelial growth. However, the supplementation of dihydroorotate to the medium did not alleviate the growth defect observed in ΔFgDHODHII mutants (Fig. 5B). Furthermore, mycelial growth in two ΔFgDHODHII mutants was restored by exogenously adding UMP, uracil or uridine under quinofumelin treatments. These findings indicate that the defects of FgDHODHII are functionally equivalent to complete inhibition of its activity by quinofumelin and further confirm that the mechanism of action for quinofumelin involves de novo synthesis of uracil while identifying FgDHODH as its specific target.

Molecular docking is a computational procedure employed to predict the optimal binding orientation of a ligand with its macromolecular target (receptor), thereby forming a stable complex. In concise terms, docking is a molecular modeling technique employed to predict the interaction between a protein (enzyme) and small molecules (ligands). In this study, we conducted molecular docking analysis between quinofumelin and the target protein FgDHODHII. The docking results unveiled a robust affinity between FgDHODHII and quinofumelin. Furthermore, these findings provided additional evidence that the strong affinity of quinofumelin with FgDHODHII disrupts its activity and function.

Surface Plasmon Resonance (SPR) technique is employed for real-time measurement of molecular interactions21. The SPR sensor map characterizes intermolecular interaction rates primarily by conveying the kinetic rate constant, while it characterizes the binding level between molecules mainly by conveying the affinity constant22-23. In this study, we used SPR technology to measure the binding energy and kinetics between FgDHODHII and quinofumelin. The binding kinetics of FgDHODHII and quinofumelin demonstrated fast-binding and fast-dissociation modes. Soluble FgDHODHII exhibited an ideal affinity (Kd) of 6.606×10−6 M with quinofumelin, falling within the suitable range of 10−3 M to 10−6 M for biomacromolecule-small molecule interactions. Furthermore, the binding affinity between FgDHODHII and quinofumelin was confirmed by a microscale thermophoresis (MST) assay. Therefore, these data strongly support the precise interaction between quinofumelin and its target protein FgDHODHII. This newly discovered selective target for antimicrobial agents offers a valuable resource for the design and development of targeted pesticides. Unlike many other dehydrogenases that require NAD or NADP as an electron acceptor, DHODH transfers electrons to ubiquinone, which then transfers electrons into respiratory complex III and eventually to oxygen24. Thus, the production of orotate in pyrimidine biosynthesis is coupled with the electron transport system and energy production, exerting additional control over nucleotide synthesis and ensuring its coupling to cellular energy metabolism. In the pyrimidine biosynthesis pathway, DHODHII regulates the conversion from dihydroorotate to orotate. Our findings demonstrated that DHODHII is target of action for quinofumelin. A previous study reported that exogenous orotate but not exogenous dihydroorotate recovered mycelial growth inhibited by quinofumelin, indicating that quinofumelin inhibits the formation of orotate in the pyrimidine biosynthesis pathway25. However, we observed that the supplementation of orotate to the medium did not alleviate the growth defect of ΔFgDHODHII mutants. Considering the disparity in mycelial growth inhibition between F. graminearum and other fungi, our data suggests that a nonclassical two-site ping-pong mechanism is the most suitable model for DHODH as previously reported14. In the classical ping-pong mechanism, DHODH interacts with dihydroorotate and ubiquinol in an identical confirmation, while orotate and ubiquinone combine with an alternative form26. However, similar to brequinar sodium inhibiting the activity of DHODH isolated from mammalian sources, FgDHODHII may follow a nonclassical, two-site ping-pong mechanism. In this nonclassical ping-pong mechanism, the two substrates may not share chemical resemblance and therefore do not bind to a single binding site but likely bind to two adjacent sites that facilitate reactant transfer. The precise nature of the interaction between quinofumelin and FgDHODHII may require further elucidation studies.

Materials and methods

Strains, culture conditions and fungicides

F. graminearum sensitive strain PH-1 and 9 wild-type strains (F. graminearum and F. asiaticum), and the PH-1 deletion mutants ΔFgDHODHII were maintained in Fungicide Biology Laboratory, Nanjing Agricultural University.

The media used in this study included potato dextrose agar (PDA) medium, Yeast extract peptone dextrose (YEPD) medium, Czapek Solution Agar (CZA) medium and LB medium.

Chemical-grade quinofumelin (Mitsui Agricultural Chemical Company of Japan) was prepared with dimethyl sulfoxide at a concentration of 1 × 104 μg/mL and stored at 4°C.

Transcriptomic assay of mycelial growth in F. graminearum as affected by quinofumelin

A total of 20 mL of YEPD medium containing 1 mL of conidia suspension (1 × 105 conidia/mL) was incubated with shaking (175 rpm/min) at 25°C. After 24 h, the medium was added with quinofumelin (EC90) at a concentration of 1 μg/mL, while an equal amount of dimethyl sulfoxide was added as the control (CK). The incubation continued for another 48 h, followed by filtration and collection of hyphae. Subsequently, Genedenovo Biotechnology Co., Ltd. (Guangzhou, China) performed sequencing on the Illumina platform. Quality control measures were implemented to remove linker sequences and low-quality bases from the original reads in order to obtain clean data readings. The filtered data underwent analysis for base composition and mass distribution to visualize data quality. Based on the gene expression data, we used R (http://www.r-project.org/) for conducting principal component analysis, while the “pheatmap” package was used to generate the clustering heatmap. DESeq2 was applied to analyze group differences based on quantitative expression results. The screening threshold of P value <0.05 and FC≥1.5 or FC≤0.67 was used to identify unigenes with significant differential expression between the two groups using a t-test. Finally, functional enrichment analysis of GO and KEGG pathways was performed on these differentially expressed unigenes.

Metabonomic assay of mycelial growth in F. graminearum as affected by quinofumelin

Transcriptome and metabolome samples were obtained from the same treatment group. The widely targeted metabolome analysis was conducted by Metware Biotechnology Co., Ltd. (Wuhan, China). The frozen sample stored at -80°C was thawed on ice. A 400 μL solution (Methanol: Water = 7:3, v/v) containing an internal standard was added to a 20 mg sample and vortexed for 3 min. Subsequently, the sample was sonicated in an ice bath for 10 min and vortexed for 1 min before being placed at -20°C for 30 min. After centrifugation at 12000 rpm for 10 min (4°C). The sediment was removed, followed by another centrifugation of the supernatant at 12000 rpm for 3 min (4°C). A volume of 200 μL aliquots of supernatant was transferred for LC-MS analysis. All samples were acquired using the LC-MS system according to machine orders. The original data file acquired by LC-MS was converted into mzML format using ProteoWizard software. Peak extraction, peak alignment, and retention time correction were performed by XCMS program.

The identified metabolites were analyzed using MBEole2.0 (https://csbg.cnb.csic.es/mbrole2/index.php) and KEGG (http://www.kegg.jp). PLS-DA analysis was used to identify the differentially accumulated metabolites (DAMs), with a screening threshold of P value<0.05 and FC≥1.5 or FC≤0.67. Each sample was replicated six times for biological consistency.

Recovery assay of quinofumelin-induced mycelial growth inhibition in F. graminearum

F. graminearum sensitive strain PH-1 and 9 wild-type strains (F. graminearum and F. asiaticum), and the PH-1 deletion mutants ΔFgDHODHII, were preincubated on a PDA plate to form mycelial colonies. Mycelial plugs (5 mm in diameter) from the edge of a 3-day-old colony were removed and inoculated onto a CZA plate containing quinofumelin (1 μg/mL) and dihydroorotate, UMP, uridine or uracil at 50 μg/mL, followed by measurement of colony diameters. Three replicates were conducted for each concentration.

Construction of deletion mutant vectors

The deletion mutant of FgDHODHII was generated through protoplast transformation, as previously described with some modifications in the transformation procedures27. For the transformation process, washed protoplasts were suspended in 160 μL STC and 40 μL SPTC (40% polyethyleneglycol 6000 in STC), along with 2 μg of the aforementioned DNA fragments and 10 μL of heparin (5 mg/mL in STC) to form protoplast suspensions. Subsequently, these mixtures were incubated on ice for 30 min, followed by mixing with 1 mL of SPTC and further incubation at room temperature for 20 min. The resulting suspensions were gently mixed into RM (100 mL) containing 50 μg/mL uridine and poured into plates. After incubating for 24 h at 25°C, the plates were overlaid with SRM containing 50 μg/mL uridine and continuously incubated for an additional 4 days. The putative transformants could be observed growing on SRM. All primers used are listed in SI Appendix, Table 1.

Molecular docking

The binding affinity of quinofumelin with FgDHODHII protein was assessed through molecular docking using AutoDockTools-1.5.7 software. The three-dimensional structure of quinofumelin (PubChem CID: 23160856) used for docking was obtained from PubChem database, while the protein structure models were sourced from the Uniport database. In the docking studies, binding energy was employed as a metric to evaluate the strength of protein-ligand interaction, where lower values indicated higher affinity between proteins and ligands. The resulting docking outcomes were visualized using PyMOL software version 2.5.4.

Surface plasmon resonance (SPR) verification of specific interactions between FgDHODHII and quinofumelin

The cDNA sequence of FgDHODHII was analyzed and codon optimization was performed for the E. coli expression system28-29. Subsequently, it was ligated into the pCold-9×his-MBP plasmid and the pET-28a plasmid at the NdeI and EcoRI sites within the multiple cloning sites to construct recombinant plasmids: pCold-9×his-MBP-TEV-DHODHII30, pET-28a (+)-DHODHII and pET-28a (+)-DHODHII124-467 (with removal of insoluble fragments) (Supporting Information, Fig. S5). Plasmid DNA was extracted using the E.Z.N.A Plasmid Extraction DNA Mini Kit I protocol (Omega, China). For protein overexpression, the exacted plasmids were separately transformed into E. coli BL21 strain31. A single clone obtained from each transformation was used for protein overexpression and purification. Upon reaching an OD600 of 1.0, 0.6 mM isopropyl β-dthiogalactopyranoside (IPTG) was added to induce protein overexpression. Following induction at 16°C for 9 h, cells were harvested, lysed, and purified using a Ni-NTA column (QIAGEN).

The interaction between FgDHODHII and quinofumelin was evaluated using Biacore T200 instrument (GE Healthcare). The CM7 sensor chip was employed for immobilizing 50 μg/mL of purified FgDHODHII protein via amino coupling method at a flow rate of 10 μL/min, utilizing 1×PBS-P+ containing 5% DMSO as running buffer. The desired coupling amount of FgDHODHII protein on the chip surface was targeted to be within the range of 12000-24000 RU. To investigate binding affinity, quinofumelin concentrations were diluted from 3.90625 to 62.5 μM in a constant flow rate running buffer containing 5% DMSO. Following sample injection, a switching buffer was introduced to allow spontaneous dissociation of quinofumelin from FgDHODHII on the chip surface, and this dissociation process was continuously monitored in real time by measuring response values. The affinity value (Kd) was determined through kinetic analysis or steady-state affinity method.

Identification of specific interactions between FgDHODHII and quinofumelin using microscale thermophoresis (MST)

To construct the FgDHODHII-GFP cassette, the FgDHODHII fragment containing the native promoter and ORF (without the stop codon) was amplified. Subsequently, the resulting PCR product was co-transformed with XhoI-digested pYF11 into E. coli DH5α as previously reported32. The fusion vector was then used to transform PH-1 to obtain the FgDHODHII-GFP strain. The GFP fusion protein was extracted from mycelia. Two Monolith NT.115 standard capillaries were employed for repeated absorption of a specific protein concentration, and placed in positions 1 and 2 of the sample tray in the MonoTemper Monolith NT.115 instrument. Sixteen PCR tubes were prepared, with 10 μL DMSO added to PCR tubes 2-16 and 20 μL of 5000 μM quinofumelin (or phenamacril) dissolved in DMSO added to PCR tube 1. Then, 10 μL of quinofumelin from PCR tube 1 was transferred and thoroughly mixed with PCR tube 2 using a pipette. These steps were repeated for PCR tubes 3–16. Following mixing with PCR tube 16 (discarding the remaining 10 μL), each PCR tube was thoroughly mixed with 190 μL DHODHII-GFP protein, and incubated at room temperature for 30 min. Monolith NT.115 standard capillaries were inserted into each PCR tube to allow solution entry into the capillary. The capillaries 1-16 were subsequently placed in the sample tray of the device and the program was initiated. All measurements were performed using 80% (Auto-detect) Nano-RED. MonoTemper Monolith NT.115 software was used to evaluate the binding affinity between the target protein and quinofumelin using the Kd values and the signal to noise ratio, and the results were combined and analyzed using Mo. Affinity Analysis v2.3 software33.

Statistical analysis

The data in the study were analyzed using analysis of variance (ANOVA) with the SPSS 14.0 software (SPSS Inc. Chicago, IL, USA). When ANOVA yielded a significant result (p = 0.05), means were compared using Fisher”s protected least significant difference (PLSD).

eLife Assessment

In this valuable study, the authors show the physiological response and molecular pathway mediating the effect of quinofumelin, a developed fungicide with an unknown mechanism. The authors present convincing data suggesting the involvement of the uridine/uracil biosynthesis pathway, by combining in vivo microbiology characterization as well as in vitro biochemical binding results.

Reviewer #1 (Public review):

Summary:

Phytophathogens including fungal pathogens such as F. graminearum remain a major threat to agriculture and food security. Several agriculturally relevant fungicides including the potent Quinofumelin have been discovered to date, yet the mechanisms of their action and specific targets within the cell remain unclear. This paper sets out to contribute to addressing these outstanding questions.

Strengths:

The paper is generally well-written and provides convincing data to support their claims for the impact of Quinofumelin on fungal growth, the target of the drug, and the potential mechanism. Critically the authors identify an important pyrimidine pathway dihydroorotate dehydrogenase (DHODH) gene FgDHODHII in the pathway or mechanism of the drug from the prominent plant pathogen F. graminearum, confirming it as the target for Quinofumelin. The evidence is supported by transcriptomic, metabolomic as well as MST, SPR, molecular docking/structural biology analyses.

Weaknesses:

Whilst the study adds to our knowledge about this drug, it is, however, worth stating that previous reports (although in different organisms) by Higashimura et al., 2022 https://pmc.ncbi.nlm.nih.gov/articles/PMC9716045/ had already identified DHODH as the target for Quinofumelin and hence this knowledge is not new and hence the authors may want to tone down the claim that they discovered this mechanism and also give sufficient credit to the previous authors work at the start of the write-up in the introduction section rather than in passing as they did with reference 25? other specific recommendations to improve the text are provided in the recommendations for authors section below.

Reviewer #2 (Public review):

Summary:

In the current study, the authors aim to identify the mode of action/molecular mechanism of characterized a fungicide, quinofumelin, and its biological impact on transcriptomics and metabolomics in Fusarium graminearum and other Fusarium species. Two sets of data were generated between quinofumelin and no treatment group, and differentially abundant transcripts and metabolites were identified. The authors further focused on uridine/uracil biosynthesis pathway, considering the significant up- and down-regulation observed in final metabolites and some of the genes in the pathways. Using a deletion mutant of one of the genes and in vitro biochemical assays, the authors concluded that quinofumelin binds to the dihydroorotate dehydrogenase.

Strengths:

Omics datasets were leveraged to understand the physiological impact of quinofumelin, showing the intracellular impact of the fungicide. The characterization of FgDHODHII deletion strains with supplemented metabolites clearly showed the impact of the enzyme on fungal growth.

Weaknesses:

Some interpretation of results is not accurate and some experiments lack controls. The comparison between quinofumelin-treated deletion strains, in the presence of different metabolites didn't suggest the fungicide is FgDHODHII specific. A wild type is required in this experiment.

Potential Impact: Confirming the target of quinofumelin may help understand its resistance mehchanism, and further development of other inhibitory molecules against the target.

The manuscript would benefit more in explaining the study rationale if more background on previous characterization of this fungicide on Fusarium is given.

Reviewer #3 (Public review):

Summary:

The manuscript shows the mechanism of action of quinofumelin, a novel fungicide, against the fungus Fusarium graminearum. Through omics analysis, phenotypic analysis, and in silico approaches, the role of quinofumelin in targeting DHODH is uncovered.

Strengths:

The phenotypic analysis and mutant generation are nice data and add to the role of metabolites in bypassing pyrimidine biosynthesis.

Weaknesses:

The role of DHODH in this class of fungicides has been known and this data does not add any further significance to the field. The work of Higashimura et al is not appreciated well enough as they already showed the role of quinofumelin upon DHODH II.

There is no mention of the other fungicide within this class ipflufenoquin, as there is ample data on this molecule.

Author response:

The following is the authors’ response to the original reviews.

Reviewer #1 (Public review):

Summary:

Phytophathogens including fungal pathogens such as F. graminearum remain a major threat to agriculture and food security. Several agriculturally relevant fungicides including the potent Quinofumelin have been discovered to date, yet the mechanisms of their action and specific targets within the cell remain unclear. This paper sets out to contribute to addressing these outstanding questions.

We appreciate the reviewer's accurate summary of our manuscript.

Strengths:

The paper is generally well-written and provides convincing data to support their claims for the impact of Quinofumelin on fungal growth, the target of the drug, and the potential mechanism. Critically the authors identify an important pyrimidine pathway dihydroorotate dehydrogenase (DHODH) gene FgDHODHII in the pathway or mechanism of the drug from the prominent plant pathogen F. graminearum, confirming it as the target for Quinofumelin. The evidence is supported by transcriptomic, metabolomic as well as MST, SPR, molecular docking/structural biology analyses.

We appreciate the reviewer's recognition of the strengths of our manuscript.

Weaknesses:

Whilst the study adds to our knowledge about this drug, it is, however, worth stating that previous reports (although in different organisms) by Higashimura et al., 2022 https://pmc.ncbi.nlm.nih.gov/articles/PMC9716045/ had already identified DHODH as the target for Quinofumelin and hence this knowledge is not new and hence the authors may want to tone down the claim that they discovered this mechanism and also give sufficient credit to the previous authors work at the start of the write-up in the introduction section rather than in passing as they did with reference 25? other specific recommendations to improve the text are provided in the recommendations for authors section below.

We appreciate the reviewer's suggestion. In the revised manuscript, we have incorporated the reference in the introduction section and expanded the discussion of previous work on quinofumelin by Higashimura et al., 2022 in the discussion section to more effectively contextualize their contributions. Moreover, we have made revisions and provided responses in accordance with the recommendations.

Reviewer #2 (Public review):

Summary:

In the current study, the authors aim to identify the mode of action/molecular mechanism of characterized a fungicide, quinofumelin, and its biological impact on transcriptomics and metabolomics in Fusarium graminearum and other Fusarium species. Two sets of data were generated between quinofumelin and no treatment group, and differentially abundant transcripts and metabolites were identified. The authors further focused on uridine/uracil biosynthesis pathway, considering the significant up- and down-regulation observed in final metabolites and some of the genes in the pathways. Using a deletion mutant of one of the genes and in vitro biochemical assays, the authors concluded that quinofumelin binds to the dihydroorotate dehydrogenase.

We appreciate the reviewer's accurate summary of our manuscript.

Strengths:

Omics datasets were leveraged to understand the physiological impact of quinofumelin, showing the intracellular impact of the fungicide. The characterization of FgDHODHII deletion strains with supplemented metabolites clearly showed the impact of the enzyme on fungal growth.

We appreciate the reviewer's recognition of the strengths of our manuscript.

Weaknesses:

Some interpretation of results is not accurate and some experiments lack controls. The comparison between quinofumelin-treated deletion strains, in the presence of different metabolites didn't suggest the fungicide is FgDHODHII specific. A wild type is required in this experiment.

Potential Impact: Confirming the target of quinofumelin may help understand its resistance mehchanism, and further development of other inhibitory molecules against the target.

The manuscript would benefit more in explaining the study rationale if more background on previous characterization of this fungicide on Fusarium is given.

We appreciate the reviewer's suggestion. Under no treatment with quinofumelin, mycelial growth remains normal and does not require restoration. In the presence of quinofumelin treatment, the supplementation of downstream metabolites in the de novo pyrimidine biosynthesis pathway can restore mycelial growth that is inhibited by quinofumelin. The wild-type control group is illustrated in Figure 4. Figure 5b depicts the phenotypes of the deletion mutants. With respect to the relationship among quinofumelin, FgDHODHII, and other metabolites, quinofumelin specifically targets the key enzyme FgDHODHII in the de novo pyrimidine biosynthesis pathway, disrupting the conversion of dihydroorotate to orotate, which consequently inhibits the synthesis downstream metabolites including uracil. In our previous study, quinofumelin not only exhibited excellent antifungal activity against the mycelial growth and spore germination of F. graminearum, but also inhibited the biosynthesis of deoxynivalenol (DON). We have added this part to the introduction section.

Reviewer #3 (Public review):

Summary:

The manuscript shows the mechanism of action of quinofumelin, a novel fungicide, against the fungus Fusarium graminearum. Through omics analysis, phenotypic analysis, and in silico approaches, the role of quinofumelin in targeting DHODH is uncovered.

We appreciate the reviewer's accurate summary of our manuscript.

Strengths:

The phenotypic analysis and mutant generation are nice data and add to the role of metabolites in bypassing pyrimidine biosynthesis.

We appreciate the reviewer's recognition of the strengths of our manuscript.

Weaknesses:

The role of DHODH in this class of fungicides has been known and this data does not add any further significance to the field. The work of Higashimura et al is not appreciated well enough as they already showed the role of quinofumelin upon DHODH II.

There is no mention of the other fungicide within this class ipflufenoquin, as there is ample data on this molecule.

We appreciate the reviewer's suggestion. We sincerely appreciate the reviewer's insightful comment regarding the work of Higashimura et al. We agree that their investigation into the role of quinofumelin in DHODH II inhibition provides critical foundational insights for this field. In the revised manuscript, we have incorporated the reference in the introduction section and expanded the discussion of their work in the discussion section to more effectively contextualize their contributions. The information regarding action mechanism of ipflufenoquin against filamentous fungi was added in discussion section.

Reviewer #1 (Recommendations for the authors):

(1) Given that the DHODH gene had been identified as a target earlier, could the authors perform blast experiments with this gene instead and let us know the percentage similarity between the FgDHODHII gene and the Pyricularia oryzae class II DHODH gene in the report by Higashimura et al., 2022.

BLAST experiment revealed that the percentage similarity between the FgDHODHII gene and the class II DHODH gene of P. oryzae was 55.41%. We have added the description ‘Additionally, the amino acid sequence of the FgDHODHII exhibits 55.41% similarity to that of DHODHII from Pyricularia oryzae, as previously reported (Higashimura et al., 2022)’ in section Results.

(2) Abstract:

The authors started abbreviating new terms e.g. DEG, DMP, etc but then all of a sudden stopped and introduced UMP with no full meaning of the abbreviation. Please give the full meaning of all abbreviations in the text, UMP, STC, RM, etc.

We have provided the full meaning for all abbreviations as requested.

(3) Introduction section:

The introduction talks very little about the work of other groups on quinofumelin. Perhaps add this information in and reference them including the work of Higashimura et al., 2022 which has done quite significant work on this topic but is not even mentioned in the background

We have added the work of other groups on quinofumelin in section introduction.

(4) General statements:

Please show a model of the pyrimidine pathway that quinofumelin attacks to make it easier for the reader to understand the context. They could just copy this from KEGG

We have added the model (Fig. 7).

(5) Line 186:

The authors did a great job of demonstrating interactions with the Quinofumelin and went to lengths to perform MST, SPR, molecular docking, and structural biology analyses yet in the end provide no details about the specific amino acid residues involved in the interaction. I would suggest that site-directed mutagenesis studies be performed on FgDHODHII to identify specific amino acid residues that interact with Quinofumelin and show that their disruption weakens Quinofumelin interaction with FgDHODHII.

Thank you for this insightful suggestion. We fully agree with the importance of elucidating the interaction mechanism. At present, we are conducting site-directed mutagenesis studies based on interaction sites from docking results and the mutation sites of FgDHODHII from the resistant mutants; however, due to the limitations in the accuracy of existing predictive models, this work remains ongoing. Additionally, we are undertaking co-crystallization experiments of FgDHODHII with quinofumelin to directly and precisely reveal their interaction pattern

.

(6) Line 76:

What is the reference or evidence for the statement 'In addition, quinofumelin exhibits no cross-resistance to currently extensively used fungicides, indicating its unique action target against phytopathogenic fungi.

If two fungicides share the same mechanism of action, they will exhibit cross resistance. Previous studies have demonstrated that quinofumelin retains effective antifungal activity against fungal strains resistant to commercial fungicides, indicating that quinofumelin does not exhibit cross-resistance with other commercially available fungicides and possesses a novel mechanism of action. Additionally, we have added the relevant inference.

(7) Line 80-82:

Again, considering the work of previous authors, this target is not newly discovered. Please consider toning down this statement 'This newly discovered selective target for antimicrobial agents provides a valuable resource for the design and development of targeted pesticides.'

We have rewritten the description of this sentence.

(8) Line 138: If the authors have identified DHODH in experimental groups (I assume in F. graminearum), what was the exact locus tag or gene name in F. graminearum, and why not just continue with this gene you identified or what is the point of doing a blast again to find the gene if the DHODH gene if it already came up in your transcriptomic or metabolic studies? This unfortunately doesn't make sense but could be explained better.

The information of FgDHODHII (gene ID: FGSG_09678) has been added. We have revised this part.

Reviewer #2 (Recommendations for the authors):

(1) Line 40:

Please add a reference.

We have added the reference

.

(2) Line 47:

Please add a reference.

We have added the reference.

(3) Line 50:

The lack of target diversity in existing fungicides doesn't necessarily serve as a reason for discovering new targets being more challenging than identifying new fungicides within existing categories, please consider adjusting the argument here. Instead, the authors can consider reasons for the lack of new targets in the field.

We have revised the description.

(4) Line 63:

Please cite your source with the new technology.

We have added the reference.

(5) Line 68:

What are you referring to for "targeted medicine", do you have a reference?

We have revised the description and the reference.

(6) Line 74:

One of the papers referred to "quinoxyfen", what are the similarities and differences between the two? Please elaborate for the readership.

Quinoxyfen, similar to quinofumelin, contains a quinoline ring structure. It inhibits mycelial growth by disrupting the MAP kinase signaling pathway in fungi (https://www.frac.info). In addition, quinoxyfen still exhibits excellent antifungal activity against the quinofumelin-resistant mutants (the findings from our group), indicating that action mechanism for quinofumelin and quinoxyfen differ.

(7) Line 84:

Please introduce why RNA-Seq was designed in the study first. What were the groups compared? How was the experiment set up? Without this background, it is hard to know why and how you did the experiment.

According to your suggestions, we have added the description in Section Results. In addition, the experimental process was described in Section Materials and methods as follows: A total of 20 mL of YEPD medium containing 1 mL of conidia suspension (1×105 conidia/mL) was incubated with shaking (175 rpm/min) at 25°C. After 24 h, the medium was added with quinofumelin at a concentration of 1 μg/mL, while an equal amount of dimethyl sulfoxide was added as the control (CK). The incubation continued for another 48 h, followed by ﬁltration and collection of hyphae. Carry out quantitative expression of genes, and then analyze the differences between groups based on the results of DESeq2 for quantitative expression.

(8) Figures:

The figure labeling is missing (Figures 1,2,3 etc). Please re-order your figure to match the text

.

The figures have been inserted.

(9) Line. 97:

"Volcano plot" is a common plot to visualize DEGs, you can directly refer to the name.

We have revised the description.

(10) Figure 1d, 1e:

Can you separate down- and up-regulated genes here? Does the count refer to gene number?

The expression information for down- and up-regulated genes is presented in Figure 1a and 1b. However, these bubble plots do not distinguish down- and up-regulated genes. Instead, they only display the significant enrichment of differentially expressed genes in specific metabolic pathways. To more clearly represent the data, we have added the detailed counts of down- and up-regulated genes for each metabolic pathway in Supplementary Table S1 and S2. Here, the term "count" refers to differentially expressed genes that fall within a certain pathway.

(11) Line 111:

Again, no reasoning or description of why and how the experiment was done here.

Based on the results of KEGG enrichment analysis, DEMs are associated with pathways such as thiamine metabolism, tryptophan metabolism, nitrogen metabolism, amino acid sugar and nucleotide sugar metabolism, pantothenic acid and CoA biosynthesis, and nucleotide sugar production compounds synthesis. To specifically investigate the metabolic pathways involved action mechanism of quinofumelin, we performed further metabolomic experiments. Therefore, we have added this description according the reviewer’s suggestions.

(12) Figure 2a:

It seems many more metabolites were reduced than increased. Is this expected? Due to the antifungal activity of this compound, how sick is the fungus upon treatment? A physiological study on F. graminearum (in a dose-dependent manner) should be done prior to the omics study. Why do you think there's a stark difference between positive and negative modes in terms of number of metabolites down- and up-regulated?

Quinofumelin demonstrates exceptional antifungal activity against Fusarium graminearum. The results indicate that the number of reduced metabolites significantly exceeds the number of increased metabolites upon quinofumelin treatment. Mycelial growth is markedly inhibited under quinofumelin exposure. Prior to conducting omics studies, we performed a series of physiological and biochemical experiments (refer to Qian Xiu's dissertation https://paper.njau.edu.cn/openfile?dbid=72&objid=50_49_57_56_49_49&flag=free). Upon quinofumelin treatment, the number of down-regulated metabolites notably surpasses that of up-regulated metabolites compared to the control group. Based on the findings from the down-regulated metabolites, we conducted experiments by exogenously supplementing these metabolites under quinofumelin treatment to investigate whether mycelial growth could be restored. The results revealed that only the exogenous addition of uracil can restore mycelial growth impaired by quinofumelin.

Quinofumelin exhibits an excellent antifungal activity against F. graminearum. At a concentration of 1 μg/mL, quinofumelin inhibits mycelial growth by up to 90%. This inhibitory effect indicates that life activities of F. graminearum are significantly disrupted by quinofumelin. Consequently, there is a marked difference in down- and up-regulated metabolites between quinofumelin-treated group and untreated control group. The detailed results were presented in Figures 1 and 2.

(13) Figure 2e:

This is a good analysis. To help represent the data more clearly, the authors can consider representing the expression using fold change with a p-value for each gene.

To more clearly represent the data, we have incorporated the information on significant differences in metabolites in the de novo pyrimidine biosynthesis pathway, as affected by quinofumelin, in accordance with the reviewer’s suggestions.

(14) Line 142:

Please indicate fold change and p-value for statistical significance. Did you validate this by RT-qPCR?

We validated the expression level of the DHODH gene under quinofumelin treatment using RT-qPCR. The results indicated that, upon treatment with the EC50 and EC90 concentrations of quinofumelin, the expression of the DHODH gene was significantly reduced by 11.91% and 33.77%, respectively (P<0.05). The corresponding results have been shown in Figure S4.

(15) Line 145:

It looks like uracil is the only metabolite differentially abundant in the samples - how did you conclude this whole pathway was impacted by the treatment?

The experiments involving the exogenous supplementation of uracil revealed that the addition of uracil could restore mycelial growth inhibited by quinofumelin. Consequently, we infer that quinofumelin disrupts the de novo pyrimidine biosynthesis pathway. In addition, as uracil is the end product of the de novo pyrimidine biosynthesis pathway, the disruption of this pathway results in a reduction in uracil levels.

(16) Figure 3:

What sequence was used as the root of the tree? Why were the species chosen? Since the BLAST query was Homo sapiens sequence, would it be good to use that as the root?

FgDHODHII sequence was used as the root of the tree. These selected fungal species represent significant plant-pathogenic fungi in agriculture production. According to your suggestion, we have removed the BLAST query of Homo sapiens in Figure 3.

(17) Figure 4:

How were the concentrations used to test chosen?

Prior to this experiment, we carried out concentration-dependent exogenous supplementation experiments. The results indicated that 50 μg/mL of uracil can fully restore mycelial growth inhibited by quinofumelin. Consequently, we chose 50 μg/mL as the testing concentration.

(18) Line 164:

Why do you hypothesize supplementing dihydroorotate would restore resistance? The metabolite seemed accumulated in the treatment condition, whereas downstream metabolites were comparable or even depleted. The DHODH gene expression was suppressed. Would accumulation of dihydroorotate be associated with growth inhibition by quinofumelin? Please include the hypothesis and rationale for the experimental setup.

DHODH regulates the conversion of dihydroorotate to orotate in the de novo pyrimidine biosynthesis pathway. The inhibition of DHODH by quinofumelin results in the accumulation of dihydroorotate and the depletion of the downstream metabolites, including UMP, uridine and uracil. Consequently, downstream metabolites were considered as positive controls, while upstream metabolite dihydroorotate served as a negative control. This design further demonstrates DHODH as action target of quinofumelin against F. graminearum. In addition, the accumulation of dihydroorotate is not associated with growth inhibition by quinofumelin; however, but the depletion of downstream metabolites in the de novo pyrimidine biosynthesis pathway is closely associated with growth inhibition by quinofumelin.

(19) Line 168:

I'm not sure if this conclusion is valid from your results in Figure 4 showing which metabolites restore growth.

o minimize the potential influence of strain-specific effects, five strains were tested in the experiments shown in Figure 4. For each strain, the first row (first column) corresponds to control condition, while second row (first column) represents treatment with 1 μg/mL of quinofumelin, which completely inhibits mycelial growth. The second row (second column) for each strain represents the supplementation with 50 μg/mL of dihydroorotate fails to restore mycelial growth inhibited by quinofumelin. In contrast, the second row (third column, fourth column, fifth colomns) for each strain demonstrated that the supplementation of 50 μg/mL of UMP, uridine and uracil, respectively, can effectively restore mycelial growth inhibited by quinofumelin.

(20) Figure 5a:

The fact you saw growth of the deletion mutant means it's not lethal. However, the growth was severely inhibited.

Our experimental results indicate that the growth of the deletion mutant is lethal. The mycelial growth observed originates from mycelial plugs that were not exposed to quinofumelin, rather than from the plates amended with quinofumelin.

(21) Figure 5b:

Would you expect different restoration of growth in the presence of quinofumelin vs. no treatment? The wild type control is missing here. Any conclusions about the relationship between quinofumelin, FgDHODHII, and other metabolites in the pathway?

Under no treatment with quinofumelin, mycelial growth remains normal and does not require restoration. In the presence of quinofumelin treatment, the supplementation of downstream metabolites in the de novo pyrimidine biosynthesis pathway can restore mycelial growth that is inhibited by quinofumelin. The wild-type control group is illustrated in Figure 4. Figure 5b depicts the phenotypes of the deletion mutants. With respect to the relationship among quinofumelin, FgDHODHII, and other metabolites, quinofumelin specifically targets the key enzyme FgDHODHII in the de novo pyrimidine biosynthesis pathway, disrupting the conversion of dihydroorotate to orotate, which consequently inhibits the synthesis downstream metabolites including uracil.

(22) Figure 6b:

Lacking positive and negative controls (known binder and non-binder). What does the Kd (in comparison to other interactions) indicate in terms of binding strength?

We tested the antifungal activities of publicly reported DHODH inhibitors (such as leflunomide and teriflunomide) against F. graminearum. The results showed that these inhibitors exhibited no significant inhibitory effects against the strain PH-1. Therefore, we lacked an effective chemical for use as a positive control in subsequent experiments. Biacore experiments offers detailed insights into molecular interactions between quinofumelin and DHODHII. As shown in Figure 6b, the left panel illustrates the time-dependent kinetic curve of quinofumelin binding to DHODHII. Within the first 60 s after quinofumelin was introduced onto the DHODHII surface, it bound to the immobilized DHODHII on the chip surface, with the response value increasing proportionally to the quinofumelin concentration. Following cessation of the injection at 60 s, quinofumelin spontaneously dissociated from the DHODHII surface, leading to a corresponding decrease in the response value. The data fitting curve presented on the right panel indicates that the affinity constant KD of quinofumelin for DHODHII is 6.606×10-6 M, which falls within the typical range of KD values (10-3 ~ 10-6 M) for protein-small molecule interaction patterns. A lower KD value indicates a stronger affinity; thus, quinofumelin exhibits strong binding affinity towards DHODHII.

Reviewer #3 (Recommendations for the authors):

The authors should add information about the other molecule within this class, ipflufenoquin, and what is known about it. There are already published data on its mode of action on DHODH and the role of pyrimidine biosynthesis.

We have added the information regarding action mechanism of ipflufenoquin against filamentous fungi in discussion section.

The work of Higashimura et al is not appreciated well enough as they already showed the role of quinofumelin upon DHODH II.

We sincerely appreciate the reviewer's insightful comment regarding the work of Higashimura et al. We agree that their investigation into the role of quinofumelin in DHODH II inhibition provides critical foundational insights for this field. In the revised manuscript, we have incorporated the reference in the introduction section and expanded the discussion of their work in the discussion section to more effectively contextualize their contributions.

It is unclear how the protein model was established and this should be included. What species is the molecule from and how was it obtained? How are they different from Fusarium?

The three-dimensional structural model of F. graminearum DHODHII protein, as predicted by AlphaFold, was obtained from the UniProt database. Additionally, a detailed description along with appropriate citations has been incorporated in the ‘Manuscript’ file.

Version 2

Abstract

Abstract

Modern fungicides have made significant contributions to crop disease management, but the development of resistant fungal strains has caused their failure in disease control. Therefore, developing fungicides with novel action mechanisms is the most effective measure to manage resistance. Quinofumelin, a novel quinoline fungicide, exhibits exceptional antifungal activity against phytopathogens. However, there is currently no available information on its mechanism of action. Here, we used transcriptome and metabolome analysis to observe a co-enrichment pattern of differentially expressed genes (DEGs) and differentially accumulated metabolites (DAMs) within pyrimidine biosynthesis pathway (PBP), identifying down-regulation of dihydroorotate dehydrogenase (DHODH). Exogenous uridine monophosphate (UMP), uridine or uracil (metabolites in PBP) successfully restored quinofumelin-induced inhibition of mycelial growth in Fusarium graminearum and Fusarium asiaticum. Additionally, the deletion of FgDHODHII was determined to be lethal; however, mycelial growth of ΔFgDHODHII mutants could be restored by adding UMP, uridine or uracil. These findings indicate that the deficiencies in FgDHODHII are functionally equivalent to complete inhibition of its activity by quinofumelin. Finally, molecular docking, surface plasmon resonance (SPR) and microscale thermophoresis (MST) results strongly support the precise interaction between quinofumelin and FgDHODHII. Collectively, these findings provide compelling evidence for the involvement of de novo uracil biosynthesis as mechanism of action for quinofumelin while identifying FgDHODHII as its specific target.

Introduction

Agrochemical fungicides have a history spanning over 3000 years in the control of plant diseases, and have undergone a development process from traditional protective fungicides with multiple action sites to modern selective fungicides with single action site (Lamberth, 2022). The introduction of the first fully systemic seed treatment agent (carboxin, a succinate dehydrogenase inhibitor) in 1966 and the launch of the first broad-spectrum foliar systemic fungicide (benomyl, a tubulin polymerization inhibitor) in 1968 represent significant milestones in the advancement of modern selective fungicides with single action site (Lamberth et al., 2021; Morton et al., 2008). Owing to their exceptional selectivity, potent activity, specific target site, and low toxicity, modern selective fungicides have gained widespread acceptance and witnessed an increasing market share year after year. However, like every coin has two sides, these types of fungicides encounter substantial challenges related to serious resistance issues when deployed in agricultural fields (Lucas et al., 2015). Numerous reports have been published on plant pathogens developing resistance to modern selective fungicides, which often result in failures or reductions in the effectiveness of chemical control against plant diseases and lead to decreased crop yields and quality (https://www.frac.info) (Duan et al., 2019; King et al., 2021; Mao et al., 2024). Therefore, it is crucial to expedite the development of new-generation selective fungicides with innovative modes of action. Based on a single fungicide target, numerous modern selective fungicides can be developed (Umetsu et al., 2020). For instance, 24 commercialized available fungicides have been created so far based on the respiratory chain complex II as the target. Similarly, another 20 commercialized fungicides have been developed targeting the respiratory chain complex III. Additionally, there are already 37 commercialized fungicides that specifically target the fungicide target C14-demethylase in ergosterol biosynthesis (https://www.frac.info). However, identifying specific targets that efficiently control plant pathogens while ensuring safety for non-target organisms remains a significant challenge. The rate of discovery for new targets lags far behind current demands.

According to the latest information released by the Fungicide Resistance Action Committee (FRAC), over 30 fungicide selective targets have been identified and acknowledged in the field of fungicide biological research spanning more than 130 years. Currently, these targets have been identified based on the successful development of fungicides and are all related proteins that are essential for pathogen growth or pathogenicity. Extensive identification and functional analysis of essential genes for growth or pathogenicity in plant pathogens have been achieved with rapid advancements in molecular biology techniques; however, there is no precedent for successfully developing commercialized fungicides targeting genes involved in growth or pathogenicity. The creation of highly active modern selective fungicides currently relies on the structural characteristics of existing targets and is accomplished through technologies such as AI drug discovery and design (AIDD), and computer aided drug design (CADD) (Yang et al., 2019). The biological functions of drug targets are highly conserved, although variations may occur at critical amino acid sites during genetic evolution. Therefore, it is crucial to explore drug targets and analyze their structural characteristics in order to develop novel fungicides with enhanced activity and selectivity. Recently, the Fungicide Biology group from Nanjing Agricultural University proposed a new concept for targeting pesticides based on the principles of targeted drugs in the medicine field (Zhong et al., 2021). The objective is to create specialized targeted pesticides that exhibit higher activity, greater selectivity, stronger specialization, lower toxicity, and environmental friendliness by leveraging the structural characteristics of pesticide targets. This approach aims to provide essential technical support for scientifically controlling crop diseases.

Quinofumelin (CAS 861647-84-9) is a novel quinoline fungicide developed by Mitsui Chemicals Co., Ltd. Japan (Ito et al., 2023). Due to their significant bioactivity, quinoline compounds and their derivatives have found extensive applications in medicine and pesticides (Tao et al., 2021; Wheeler et al., 2003). In our previous study, quinofumelin not only exhibited excellent antifungal activity against the mycelial growth and spore germination of F. graminearum, but also inhibited the biosynthesis of deoxynivalenol (DON) (Xiu et al., 2021). In addition, quinofumelin exhibits no cross-resistance to currently extensively used fungicides, indicating its unique action target against phytopathogenic fungi (Tao et al., 2021). In previous studies, quinofumelin was identified as a specific inhibitor of dihydroorotate dehydrogenase class II (DHODHII) in Pyricularia oryzae, suppressing fungal growth by blocking the conversion of dihydroorotate to orotate (Higashimura et al., 2022). Gene disruption of DHODHII (PoPYR4) further confirmed its critical role in fungal pathogenicity and validated it as the action target site of quinofumelin. However, previous investigations did not report any detailed methodologies for target identification. Additionally, there was a lack of information regarding the precise interaction between quinofumelin and DHODHII. In this study, we used transcriptomic and metabolomic analyses to investigate the co-enrichment patterns of differentially expressed genes (DEGs) and differentially accumulated metabolites (DAMs). This approach enabled the identification of potential metabolic pathways were identified. Ultimately, the pyrimidine biosynthesis pathway was identified as the target. Therefore, this study unveils that the mechanism of action of quinofumelin involves de novo synthesis of uracil, thereby identifing DHODH as its specific target. As a selective fungicidal target, DHODH offers a valuable resource for the design and development of novel fungicides.

Results

Transcriptomic analysis of F. graminearum as affected by quinofumelin

RNA-seq experiments have emerged as the standard approach for quantifying and comparing gene expression levels across a wide range of species and conditions, providing critical insights into cellular phenomena (Pertea et al., 2016). To investigate the impact of quinofumelin on gene expression levels in F. graminearum, we performed an RNA-Seq experiment. For quality assessment of the RNA-Seq data, we conducted a comprehensive quality control analysis. In the principal component analysis, PC1 and PC2 accounted for 45.59% and 16.98% of the total variance, respectively. The distinct separation observed between the groups indicates significant differences in gene expression between the controls and experimental samples, which could be used for subsequent investigations (Supplementary Information, Fig. S1). The clustering heat map effectively categorized the controls and experimental samples into distinct clusters based on their gene expression profiles, demonstrating robust biological replication with minimal intra-group variability and ensuring the reliability of the data. Moreover, notable inter-group dissimilarities highlighted variations among different treatment samples (Fig. 1c).

Analysis of transcriptome sequencing data in F. graminearum PH-1 treated with quinofumelin.

a Bar chart illustrating the comparison of differentially expressed genes between EC90 treatments and the control.

b Volcano plot depicting up-regulated DEGs as red scattered dots and down-regulated DEGs as blue scattered dots.

c Cluster heat maps displaying gene relative expression values using RNA-seq data scale-standardized values.

d Bubble map presenting the results of GO enrichment analysis for DEGs.

e Bubble map showcasing the findings of KEGG enrichment analysis for DEGs.

The differential analysis unveiled a total of 234 significantly differentially expressed genes (DEGs) between the control and experimental samples. Among these findings, there were identified as up-regulated for a count of 97 genes while down-regulation was observed in case of another set comprising around 137 genes (Fig. 1a). A graphical representation was employed to depict the statistical distribution pattern exhibited by these DEGs (Fig. 1b). To gain deeper insights into gene functionalities, Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) enrichment analyses were performed on these DEGs. The results showed that they are notably enriched in seventeen distinct GO terms spanning across two major categories; eleven molecular functions along with six processes (Fig. 1d). Furthermore, the enrichment analysis highlighted significant associations with various functional attributes including transition metal ion binding, oxidoreductase activity, modified amino acid binding, amide binding, monooxygenase activity, and heme binding. Moreover, the enriched biological processes primarily involved transcriptional regulation driven by nucleic-enabled mechanisms as well as monocarboxylic acid metabolism and RNA biosynthesis. Additionally, the KEGG enrichment analysis identified specific pathways associated with thiamine metabolism, tryptophan metabolism, nitrogen metabolism, amino acid sugar and nucleotide sugar metabolism, pantothenic acid and CoA biosynthesis, as well as nucleotide sugar production compound synthesis (Fig. 1e). In the GO and KEGG enrichment analyses, the detailed counts of down- and up-regulated DEGs for each metabolic pathway are provided in Supplementary Table S1 and S2.

Metabolomic analysis of F. graminearum as affected by quinofumelin

According to the findings from the KEGG enrichment analysis of DEGs derived from transcriptomic data, these DEGs are associated with metabolism pathways as above described. To specifically investigate the metabolic pathways involved action mechanism of quinofumelin, we performed further metabolomic experiments. In metabolomics principal component analysis, the contribution rate of PC1 was determined to be 76.6% in positive ion mode and 51.2% in negative ion mode, respectively. Moreover, the quality control (QC) samples exhibited highly concentrated results, indicating excellent instrument stability and remarkable data repeatability. Furthermore, a distinct separation between groups was observed, particularly prominent in the positive ion mode, indicating significant variations in metabolites among different groups that warrant further investigation (Supplementary Information, Fig. S2a and S2b). The findings from Pearson correlation analysis were consistent with those obtained from principal component analysis (Supplementary Information, Fig. S2c and S2d). Strong correlations were identified within samples belonging to the same group, while weak correlations were observed between different groups.

Metabolomics analysis revealed a total of 713 differential metabolites, with 14 up-regulated and 699 down-regulated in the positive ion mode, as well as 230 differential metabolites, including 103 up-regulated and 127 down-regulated in the negative ion mode (Supplementary Information, Fig. 2a and Fig. S3a). Based on the classification of the identified 943 different metabolites, it was observed that amino acids and their derivatives accounted for the highest proportion (17%), followed by organic acids and their derivatives (11%), benzene and substituted derivatives (11%), aldehydes, ketones, esters (11%), and heterocyclic compounds (10%) (Fig. 2c). The volcano plot visually represents the distribution of statistical differences in metabolite abundance between the two groups (Supplementary Information, Fig. 2b and Fig. S3b).

Difference analysis and enrichment analysis of metabolome data in F. graminearum PH-1 treated with quinofumelin.

a Bar chart illustrating the number of differential accumulation metabolites (DAMs) quantity in experimental groups (EC90) and control groups (CK) in positive ion mode.

b Volcanic map of all metabolites in positive ion mode. The red scatter points represent up-regulated DAMs, while the blue scatter points represent down-regulated DAMs.

c Classification and proportion of DAMs.

d Bubble diagram of differential accumulation metabolite KEGG enrichment pathway.

e Pyrimidine de novo biosynthesis pathway in Fusarium graminearum. The left box plots represent the relative content of 8 metabolite samples, while the right box represents the relative content of 6 gene samples. The color gradient from navy blue to firebrick red indicates a progression from low to high content.

KEGG pathways enriched by differential metabolites primarily include arachidonic acid metabolism, pyrimidine metabolism, alanine, aspartate and glutamate metabolism, purine metabolism, citrate cycle, arginine and proline metabolism, fructose and mannose metabolism, nitrogen metabolism, and metabolic pathways (Fig. 2d). The classification of enrichment pathways for differential metabolites revealed that the majority of these metabolites were associated with nucleotide metabolism (25%), amino acid metabolism (25%), carbohydrate metabolism (25%), and energy metabolism (12.5%) (Supplementary Information, Fig. S3c).

Conjoint analysis of transcriptome and metabolome integration

Through transcriptome and metabolome data analysis, we observed a co-enrichment of differentially expressed genes and differentially accumulated metabolites in the pyrimidine metabolism pathway. In the expression calorigrams of metabolites and genes depicting the de novo biosynthesis pathway of pyrimidine, we identified down-regulation of dihydroorotate dehydrogenase (DHODH) in the experimental groups. Through in vitro RT-qPCR verification, it was found that after treating with quinofumelin at EC50 (0.035 μg/mL) and EC90 (1 μg/mL) concentrations, the expression of the DHODH gene was respectively decreased by 11.91% and 33.77% (P < 0.05) (Supplementary Information, Fig. S4). This down-regulation hinders the conversion of dihydroorotate to orotate, thereby inhibiting the biosynthesis of downstream metabolites orotidine-5P and uracil, leading to their decreased abundance (Fig. 2e). Thus, we conclude that the action mechanism of quinofumelin may involve inhibition of the de novo biosynthesis pathway of pyrimidine.

Identification of DHODHII in F. graminearum

The FgDHODHII gene (FGSG_09678) was identified through the integrated analysis of transcriptome and metabolome data. The gene sequence for FgDHODHII was retrieved from the National Center for Biotechnology Information (NCBI) GenBank database. FgDHODHII is a gene consisting of 1419 bp without introns and is predicted to encode a protein consisting of 472 amino acids. Orthologs of FgDHODHII were obtained using the BLAST algorithm. The phylogenetic tree was constructed based on the amino acid sequences of DHODHII homologous protein with Mega X using the neighbor-joining method. Motif pattern information was generated using the MEME suite (Bailey et al., 2009), while functional domain information and protein coordinates were obtained through CD-search (Marchler-Bauer et al., 2009). TBtools software was used for visualizing the tree representation (Chen et al., 2020). All DHODHII proteins possess the dihydroorotate dehydrogenase domain (Fig. 3), indicating high conservation among fungi. Additionally, the amino acid sequence of the FgDHODHII exhibits 55.41% similarity to that of DHODHII from Pyricularia oryzae, as previously reported (Higashimura et al., 2022).

Phylogenic tree of DHODHII proteins.

The phylogenetic tree was constructed based on the amino acid sequences of DHODHII homologous protein with Mega X using the neighbor-joining method. The bootstrap values from 1000 replications are indicated on the branches. Motif pattern information was generated using the MEME suite12, while functional domain information and protein coordinates were obtained through CD-search. The final phylogenetic tree was visualized using TBtools. FgDHODHII was highlighted in red. The amino acid sequence of DHODHII from Alternaria alternata (AA0117_g542), Aspergillus nidulans (ANIA_05909), Aspergillus oryzae (AO1008_05965), Blumeria graminis (BLGH_00861), Botrytis cinerea (Bcin15g04150), Botrytis elliptica (BELL_0162g00150), Colletotrichum asianum (GQ607_007053), Colletotrichum fioriniae (CFIO01_03831), Colletotrichum higginsianum (CH063_10143), Corynespora cassiicola (BS50DRAFT_576390), Erysiphe necator (EV44_g2092), Erysiphe pulchra (EPUL_004094), Fusarium euwallaceae (BHE90_002825), Fusarium floridanum (CEP51_012005), Fusarium fujikuroi (FFMR_05040), Fusarium graminearum (FGSG_09678), Fusarium nygamai (FNYG_10990), Fusarium odoratissimum (FOIG_08270), Fusarium oxysporum (FOXG_08390), Fusarium pseudograminearum (FPSE_04283), Fusarium solani (NechaG15121), Fusarium verticillioides (FVEG_06288), Magnaporthe oryzae (MGG_08814), Neurospora crassa (NCU06532), Oidium neolycopersici (OnM2_089032), Pestalotiopsis fici (PFICI_04984), Podospora anserina (PODANS_6_7920), Puccinia graminis (PGTG_16739), Puccinia triticina (PTTG_02070), Rhizoctonia solani (RSOLAG22IIIB_08327), Schizosaccharomyces pombe (SPAC57A10.12c), Ustilaginoidea virens (UVI_02052880), Verticillium alfalfae (VDBG_01322), Verticillium dahliae (VDAG_07935), and Zymoseptoria tritici (Mycgr3G72548) were accessed in EnsemblFungi database.

Recovery test of mycelial growth suppressed by quinofumelin in Fusarium genus

The mycelial growth of F. graminearum PH-1 was significantly inhibited on CZA plates containing quinofumelin at concentrations of 1 μg/mL (Fig. 4). To investigate the potential recovery of mycelial growth by exogenous dihydroorotate, uridine monophosphate (UMP), uridine, or uracil, CZA plates supplemented with quinofumelin were treated with these compounds. Despite the addition of exogenous dihydroorotate at 50 μg/mL failing to restore mycelial growth, the addition of exogenous UMP, uridine or uracil at 50 μg/mL successfully recovered quinofumelin-induced inhibition of mycelial growth. Similar results were observed in other strains of F. graminearum or F. asiaticum (Fig. 4 and Fig. S5). Thus, it can be concluded that quinofumelin effectively inhibits the pyrimidine biosynthesis pathways.

Recovery test of mycelial growth suppressed by quinofumelin.

all strains were incubated on CZA plates at 25°C for 3 days. F. graminearum strains PH-1, 512E-8, SX2117, XY-4, and R10. The left image shows the colony morphology, while the right image is a bar chart of colony diameters.

The absence of FgDHODHII is lethal for mycelial growth in F. graminearum

To further investigate the biological functions of FgDHODHII in F. graminearum, we employed a homologous recombination strategy to generate target gene deletion mutants. Through PCR analysis, two putative FgDHODHII deletion mutants were successfully identified. The ΔFgDHODHII mutants exhibited distinct phenotypic differences compared to the wild-type progenitor PH-1 and the respective complemented strain (Fig. 5a). These mutants displayed an inability to grow on CZA plate; however, their mycelial growth was restored when supplemented with 50 μg/mL of exogenous UMP, uridine or uracil, while no recovery was observed with 50 μg/mL of exogenous dihydroorotate (Fig. 5b).

Mycelial growth of FgDHODHII deletion mutants on CZA plates.

a Mycelial growth of the FgDHODHII deletion mutants and the parental strain was observed in the presence or absence of 50 μg/mL uridine.

b Recovery of mycelial growth of the FgDHODHII deletion mutants was observed in the presence of 50 μg/mL dihydroorotate, UMP, uridine or uracil. All strains were incubated at 25°C for 3 days.

Molecular docking

To facilitate the molecular docking process between FgDHODHII and quinofumelin, ten distinct binding models were generated using AutoDockTools-1.5.7 software, each accompanied by its respective binding energies calculation results. The most favorable conformational arrangement was determined based on its relative ranking in terms of the computed binding energy values and subsequently visualized through PyMOL 2.5.4 software (Fig. 6a). Notably, the obtained docking results revealed a robust interaction between FgDHODHII protein and quinofumelin as evidenced by a calculated negative free energy value of −6.8 kcal/mol.

Validation and analysis of quinofumelin’s binding affinity to FgDHODHII.

a Three-dimensional model of the wild-type FgDHODHII are depicted, with an enlarged view highlighting the binding site for quinofumelin on the corresponding FgDHODHII models.

b The SPR response (RU) analysis and fitting results demonstrate the interaction between quinofumelin and FgDHODHII.

c Microscale thermophoresis (MST) analysis to detect the binding of quinofumelin and FgDHODHII.

The precise interaction between FgDHODHII and quinofumelin was verified by Surface plasmon resonance analysis

Surface plasmon resonance (SPR) was utilized to evaluate the interaction affinity between quinofumelin and FgDHODHII in this study. The experiment was conducted using a BIAcore T200 (Cytiva) instrument for biomolecular interaction analysis. FgDHODHII was immobilized as the ligand on the chip surface, while quinofumelin continuously flowed through as the analyte solution. Dilute quinofumelin to different concentration gradients (62.5, 32.15, 15.63, 7.81, 3.91 μM), measure the binding energy, and then fit the binding equilibrium dissociation constant (Kd) value (Fig. 6b). The SPR results revealed that quinofumelin exhibited fast-binding and fast-dissociation modes when interacting with the target protein FgDHODHII. The calculated KD value was determined as 6.606×10−6 M, indicating specific binding of quinofumelin to FgDHODHII.

The interaction between FgDHODHII and quinofumelin was confirmed by microscale thermophoresis

In this study, microscale thermophoresis (MST) was further used to evaluate the interaction affinity between FgDHODHII and quinofumelin. The strength of the binding can be inferred from the Kd value. The experiment was conducted using MO.Control v1.6.1 and data analysis was performed using MO. Affinity Analysis v2.3. By introducing a fluorescent label GFP to the FgDHODHII protein while maintaining its concentration constant, non-fluorescent labeled small molecules were diluted into 16 concentration gradients. Subsequently, the protein solution and small molecule diluent were mixed at a ratio of 19:1 before being loaded into the capillary for further analysis in the machine. The MST assay demonstrated that FgDHODHII exhibited specific binding to quinofumelin in vitro, with a dissociation constant (Kd) of 0.471 ± 0.268 μM, whereas no binding was observed with phenamacril which lacks cross-resistance with quinofumelin (Fig. 6c).

A model of the pyrimidine pathway that quinofumelin attacks.

Discussion

In over 130 years of research on fungicide biology, more than 30 novel targets for selective fungicides have been identified. Based on these newly discovered targets, over 200 modern selective fungicides have been developed. Consequently, the identification of novel targets for fungicides poses a significantly greater challenge compared to the development of new fungicides. Currently, modern selective fungicides with single action site are widely applied in agricultural practices for crop disease management. However, these specific types of fungicides encounter substantial obstacles associated with resistance issues when deployed in agricultural fields. Numerous studies have extensively documented cases of plant pathogens developing resistance to modern selective fungicides, resulting in failures or reductions in their efficacy against plant diseases and subsequently leading to decreased crop yields and quality (Duan et al., 2019; King et al., 2021; Mao et al., 2024). For instance, Botrytis cinerea, a significant plant pathogenic fungus, has emerged as multi-resistant population that is resistant to all currently used modern selective fungicides (Shao et al., 2020; Weber et al., 2023). This poses a formidable challenge for the control of crop grey mold disease caused by B. cinerea. Therefore, there is an urgent need for the development of new fungicides with novel mechanisms of action to effectively manage multiple resistance.

The novel quinoline, quinofumelin, developed by Mitsui Chemicals Co., Ltd. Japan, has demonstrated remarkable antifungal activity against plant pathogens such as F. graminearum and Sclerotinia sclerotiorum (Tao et al., 2021; Xiu et al., 2021). Moreover, our research group found that quinofumelin exhibits no cross-resistance to all types of currently used modern selective fungicides (Tao et al., 2021). These findings indicate that quinofumelin possesses a distinctive mode of action against plant pathogens. Quinofumelin has been identified as a agrochemical fungicide targeting DHODHII in Pyricularia oryzae (Higashimura et al., 2022). Similarly, ipflufenoquin, another agrochemical fungicide with an analogous mechanism of action to quinofumelin, has been reported to target DHODH in Aspergillus fumigatus (Rhijn et al., 2023). To elucidate the mode of action of quinofumelin against F. graminearum, we conducted a comprehensive investigation using integrated transcriptomic and metabolomic analyses. The transcriptome data revealed that DEGs were involved in ‘tryptophan metabolism’ and ‘amino sugar and nucleotide sugar metabolism’. Furthermore, the metabolome data identified DAMs enriched in ‘nucleotide metabolism’ and ‘amino acid metabolism’. Interestingly, we observed a co-enrichment of DEGs and DAMs in the pyrimidine metabolism pathway. Specifically, we identified a down-regulation of dihydroorotate dehydrogenase (DHODH) in the experimental groups, which impeded the conversion of dihydroorotate to orotate. Consequently, this inhibition led to decreased biosynthesis of downstream metabolites orotidine-5P and uracil. Therefore, we conducted an experiment using F. graminearum standard strain PH-1 to quantify mycelial growth by supplementing exogenous uracil, which is the end product of the uracil biosynthesis pathway. This was performed in the presence of complete inhibition of mycelial growth caused by quinofumelin. Surprisingly, the addition of exogenous uracil restored the inhibited mycelial growth induced by quinofumelin (Fig. 4). Meanwhile, similar results were successfully verified with four other F. graminearum strains and five F. asiaticum strains (Fig. 4 and Fig. S4). Based on our depicted pyrimidine biosynthesis pathway (Fig. 2E), we also supplemented other intermediate products such as dihydroorotate, orotate, UMP, and uridine externally. Except for dihydroorotate and orotate, supplementation of exogenous UMP and uridine effectively restored mycelial growth inhibited by quinofumelin as well. These findings strongly support that quinofumelin acts through modulation of the uracil biosynthesis pathway. Based on these findings and in conjunction with previous studies ((Higashimura et al., 2022; Rhijn et al., 2023), these agrochemical fungicides are categorized under FRAC group 52. This classification targets dihydroorotate dehydrogenase (DHODH) during de novo pyrimidine biosynthesis pathway, as outlined by the Fungicide Resistance Action Committee (https://www.frac.info/).As a key enzyme in the de novo pyrimidine biosynthesis pathway, DHODH has previously been investigated as a potential target for inhibiting infectious organisms and as an antitumor agent in mammals (Chen et al., 1992; Mao et al., 2021). Inhibiting DHODH can effectively impede the synthesis of DNA, RNA, and glycoproteins, thereby regulating aberrant cell metabolism and proliferation to ultimately achieve therapeutic objectives (Boukalova et al., 2020). Within mammalian cells, DHODH is closely associated with the mitochondrial membrane and plays a pivotal role in the respiratory chain through its interaction with ubiquinone (Chen et al., 1976). Here, using amino acid sequences from Homo sapiens DHODHII as a reference query, we have identified FgDHODHII (FGSG_09678), a gene comprising of 1419 bp without any introns, which encodes 472 amino acids in F. graminearum. FgDHODHII possesses a typical DHOD-2 like domain (NCBI: cd04738) and belongs to class 2 DHODs. Class 2 DHODHs are membrane-bound enzymes, whereas class 1 enzymes are located in the cytosol (Nørager et al., 2002). The DHODH enzyme catalyzes a unique redox reaction in the de novo pyrimidine biosynthesis pathway(Dorasamy et al., 2017). The DHOD-2 like domain can bind orotate and flavin mononucleotide (FMN), facilitating the oxidation of dihydroorotate to orotate and the reduction of the FMN cofactor (Alves et al., 2015). Based on the aforementioned findings, we have concluded that DHODHII functions as the target of action for quinofumelin against F. graminearum. Consequently, we performed a protoplast transformation assay to delete FgDHODHII in F. graminearum and further validate the precise target of action for quinofumelin. Typically, most fungicide targets are either lethal genes or essential pathogenic genes. Therefore, it is plausible that the inactivation of FgDHODHII may be lethal. In our genetic transformation experiment, exogenous uridine was supplemented to successfully obtain deletion mutants of the FgDHODHII gene. Fortunately, two ΔFgDHODHII mutants were successfully obtained. These mutants exhibited an inability to grow on PDA or CZA medium; however, mycelial growth was fully restored upon addition of uridine to the medium (Fig. 5A). Additionally, exogenous supplementation of UMP or uracil effectively restored mycelial growth. However, the supplementation of dihydroorotate to the medium did not alleviate the growth defect observed in ΔFgDHODHII mutants (Fig. 5B). Furthermore, mycelial growth in two ΔFgDHODHII mutants was restored by exogenously adding UMP, uracil or uridine under quinofumelin treatments. These findings indicate that the defects of FgDHODHII are functionally equivalent to complete inhibition of its activity by quinofumelin and further confirm that the mechanism of action for quinofumelin involves de novo synthesis of uracil while identifying FgDHODHII as its specific target.

Molecular docking is a computational procedure employed to predict the optimal binding orientation of a ligand with its macromolecular target (receptor), thereby forming a stable complex. In concise terms, docking is a molecular modeling technique employed to predict the interaction between a protein (enzyme) and small molecules (ligands). In this study, we conducted molecular docking analysis between quinofumelin and the target protein FgDHODHII. The docking results unveiled a robust affinity between FgDHODHII and quinofumelin. Furthermore, these findings provided additional evidence that the strong affinity of quinofumelin with FgDHODHII disrupts its activity and function.

Surface Plasmon Resonance (SPR) technique is employed for real-time measurement of molecular interactions (Olaru et al., 2015). The SPR sensor map characterizes intermolecular interaction rates primarily by conveying the kinetic rate constant, while it characterizes the binding level between molecules mainly by conveying the affinity constant (Guo, 2012; Piliarik et al., 2009). In this study, we used SPR technology to measure the binding energy and kinetics between FgDHODHII and quinofumelin. The binding kinetics of FgDHODHII and quinofumelin demonstrated fast-binding and fast-dissociation modes. Soluble FgDHODHII exhibited an ideal affinity (Kd) of 6.606×10−6 M with quinofumelin, falling within the suitable range of 10−3 M to 10− 6 M for biomacromolecule-small molecule interactions. Furthermore, the binding affinity between FgDHODHII and quinofumelin was confirmed by a microscale thermophoresis (MST) assay. Therefore, these data strongly support the precise interaction between quinofumelin and its target protein FgDHODHII. This newly discovered selective target for antimicrobial agents offers a valuable resource for the design and development of targeted pesticides.

Unlike many other dehydrogenases that require NAD or NADP as an electron acceptor, DHODH transfers electrons to ubiquinone, which then transfers electrons into respiratory complex III and eventually to oxygen (Wang et al., 2016). Thus, the production of orotate in pyrimidine biosynthesis is coupled with the electron transport system and energy production, exerting additional control over nucleotide synthesis and ensuring its coupling to cellular energy metabolism. In the pyrimidine biosynthesis pathway, DHODHII regulates the conversion from dihydroorotate to orotate. Our findings demonstrated that DHODHII is target of action for quinofumelin. A previous study reported that exogenous orotate but not exogenous dihydroorotate recovered mycelial growth inhibited by quinofumelin, indicating that quinofumelin inhibits the formation of orotate in the pyrimidine biosynthesis pathway (Higashimura et al., 2022). However, we observed that the supplementation of orotate to the medium did not alleviate the growth defect of ΔFgDHODHII mutants. Considering the disparity in mycelial growth inhibition between F. graminearum and other fungi, our data suggests that a nonclassical two-site ping-pong mechanism is the most suitable model for DHODH as previously reported (Chen et al., 1992). In the classical ping-pong mechanism, DHODH interacts with dihydroorotate and ubiquinol in an identical confirmation, while orotate and ubiquinone combine with an alternative form (Cleland, 1963). However, similar to brequinar sodium inhibiting the activity of DHODH isolated from mammalian sources, FgDHODHII may follow a nonclassical, two-site ping-pong mechanism. In this nonclassical ping-pong mechanism, the two substrates may not share chemical resemblance and therefore do not bind to a single binding site but likely bind to two adjacent sites that facilitate reactant transfer. The precise nature of the interaction between quinofumelin and FgDHODHII may require further elucidation studies.

Materials and methods

Strains, culture conditions and fungicides

F. graminearum sensitive strain PH-1 and 9 wild-type strains (F. graminearum and F. asiaticum), and the PH-1 deletion mutants ΔFgDHODHII were maintained in Fungicide Biology Laboratory, Nanjing Agricultural University.

The media used in this study included potato dextrose agar (PDA) medium, Yeast extract peptone dextrose (YEPD) medium, Czapek Solution Agar (CZA) medium and LB medium.

Chemical-grade quinofumelin (Mitsui Agricultural Chemical Company of Japan) was prepared with dimethyl sulfoxide at a concentration of 1 × 104 μg/mL and stored at 4°C.

Transcriptomic assay of mycelial growth in F. graminearum as affected by quinofumelin

A total of 20 mL of YEPD medium containing 1 mL of conidia suspension (1 × 105 conidia/mL) was incubated with shaking (175 rpm/min) at 25°C. After 24 h, the medium was added with quinofumelin (EC90) at a concentration of 1 μg/mL, while an equal amount of dimethyl sulfoxide was added as the control (CK). The incubation continued for another 48 h, followed by filtration and collection of hyphae. Subsequently, Genedenovo Biotechnology Co., Ltd. (Guangzhou, China) performed sequencing on the Illumina platform. Quality control measures were implemented to remove linker sequences and low-quality bases from the original reads in order to obtain clean data readings. The filtered data underwent analysis for base composition and mass distribution to visualize data quality. Based on the gene expression data, we used R (http://www.r-project.org/) for conducting principal component analysis, while the “pheatmap” package was used to generate the clustering heatmap. DESeq2 was applied to analyze group differences based on quantitative expression results. The screening threshold of P value <0.05 and FC≥1.5 or FC≤0.67 was used to identify unigenes with significant differential expression between the two groups using a t-test. Finally, functional enrichment analysis of GO and KEGG pathways was performed on these differentially expressed unigenes.

Metabonomic assay of mycelial growth in F. graminearum as affected by quinofumelin

Transcriptome and metabolome samples were obtained from the same treatment group. The widely targeted metabolome analysis was conducted by Metware Biotechnology Co., Ltd. (Wuhan, China). The frozen sample stored at −80°C was thawed on ice. A 400 μL solution (Methanol: Water = 7:3, v/v) containing an internal standard was added to a 20 mg sample and vortexed for 3 min. Subsequently, the sample was sonicated in an ice bath for 10 min and vortexed for 1 min before being placed at −20°C for 30 min. After centrifugation at 12000 rpm for 10 min (4°C). The sediment was removed, followed by another centrifugation of the supernatant at 12000 rpm for 3 min (4°C). A volume of 200 μL aliquots of supernatant was transferred for LC-MS analysis. All samples were acquired using the LC-MS system according to machine orders. The original data file acquired by LC-MS was converted into mzML format using ProteoWizard software. Peak extraction, peak alignment, and retention time correction were performed by XCMS program.

The identified metabolites were analyzed using MBEole2.0 (https://csbg.cnb.csic.es/mbrole2/index.php) and KEGG (http://www.kegg.jp). PLS-DA analysis was used to identify the differentially accumulated metabolites (DAMs), with a screening threshold of P value<0.05 and FC≥1.5 or FC≤0.67. Each sample was replicated six times for biological consistency.

Recovery assay of quinofumelin-induced mycelial growth inhibition in F. graminearum

F. graminearum sensitive strain PH-1 and 9 wild-type strains (F. graminearum and F. asiaticum), and the PH-1 deletion mutants ΔFgDHODHII, were preincubated on a PDA plate to form mycelial colonies. Mycelial plugs (5 mm in diameter) from the edge of a 3-day-old colony were removed and inoculated onto a CZA plate containing quinofumelin (1 μg/mL) and dihydroorotate, UMP, uridine or uracil at 50 μg/mL, followed by measurement of colony diameters. Three replicates were conducted for each concentration.

Construction of deletion mutant vectors

The deletion mutant of FgDHODHII was generated through protoplast transformation, as previously described with some modifications in the transformation procedures(Zhang et al., 2016). For the transformation process, washed protoplasts were suspended in 160 μL sorbitol-tris-calcium chloride (STC) solution and 40 μL sorbitol-peg-tris-calcium chloride (SPTC) solution (40% polyethylene glycol 6000 in STC), along with 2 μg of the aforementioned DNA fragments and 10 μL of heparin (5 mg/mL in STC) to form protoplast suspensions. Subsequently, these mixtures were incubated on ice for 30 min, followed by mixing with 1 mL of SPTC and further incubation at room temperature for 20 min. The resulting suspensions were gently mixed into regeneration medium (RM) (100 mL) containing 50 μg/mL uridine and poured into plates. After incubating for 24 h at 25°C, the plates were overlaid with selective regeneration medium (SRM) containing 50 μg/mL uridine and continuously incubated for an additional 4 days. The putative transformants could be observed growing on SRM. All primers used are listed in Table 1.

Molecular docking

The binding affinity of quinofumelin with FgDHODHII protein was assessed through molecular docking using AutoDockTools-1.5.7 software. The three-dimensional structure of quinofumelin (PubChem CID: 23160856) used for docking was obtained from PubChem database, while the three-dimensional structural model of F. graminearum DHODHII protein predicted by AlphaFold (Jumper et al., 2021) was obtained from the UniProt database (Bateman et al., 2019). In the docking studies, binding energy was employed as a metric to evaluate the strength of protein-ligand interaction, where lower values indicated higher affinity between proteins and ligands. The resulting docking outcomes were visualized using PyMOL software version 2.5.4.

Surface plasmon resonance (SPR) verification of specific interactions between FgDHODHII and quinofumelin

The cDNA sequence of FgDHODHII was analyzed and codon optimization was performed for the E. coli expression system (Kyriakopoulos et al., 2013; Yang et al., 2021). Subsequently, it was ligated into the pCold-9×his-MBP plasmid and the pET-28a plasmid at the NdeI and EcoRI sites within the multiple cloning sites to construct recombinant plasmids: pCold-9×his-MBP-TEV-DHODHII (Fang et al., 2018), pET-28a (+)-DHODHII and pET-28a (+)-DHODHII124-467 (with removal of insoluble fragments) (Supplementary Information, Fig. S6).

Plasmid DNA was extracted using the E.Z.N.A Plasmid Extraction DNA Mini Kit I protocol (Omega, China). For protein overexpression, the exacted plasmids were separately transformed into E. coli BL21 strain (Zhu et al., 2014). A single clone obtained from each transformation was used for protein overexpression and purification. Upon reaching an OD600 of 1.0, 0.6 mM isopropyl-β-D-thiogalactopyranoside (IPTG) was added to induce protein overexpression. Following induction at 16°C for 9 h, cells were harvested, lysed, and purified using a Ni-NTA column (QIAGEN).

The interaction between FgDHODHII and quinofumelin was evaluated using Biacore T200 instrument (GE Healthcare). The CM7 sensor chip was employed for immobilizing 50 μg/mL of purified FgDHODHII protein via amino coupling method at a flow rate of 10 μL/min, utilizing 1×PBS-P+ containing 5% DMSO as running buffer. The desired coupling amount of FgDHODHII protein on the chip surface was targeted to be within the range of 12000-24000 RU. To investigate binding affinity, quinofumelin concentrations were diluted from 3.90625 to 62.5 μM in a constant flow rate running buffer containing 5% DMSO. Following sample injection, a switching buffer was introduced to allow spontaneous dissociation of quinofumelin from FgDHODHII on the chip surface, and this dissociation process was continuously monitored in real time by measuring response values. The affinity value (Kd) was determined through kinetic analysis or steady-state affinity method.

Identification of specific interactions between FgDHODHII and quinofumelin using microscale thermophoresis (MST)

To construct the FgDHODHII-GFP cassette, the FgDHODHII fragment containing the native promoter and ORF (without the stop codon) was amplified. Subsequently, the resulting PCR product was co-transformed with XhoI-digested pYF11 into E. coli DH5α as previously reported (Tang et al., 2020). The fusion vector was then used to transform PH-1 to obtain the FgDHODHII-GFP strain. The GFP fusion protein was extracted from mycelia. Two Monolith NT.115 standard capillaries were employed for repeated absorption of a specific protein concentration, and placed in positions 1 and 2 of the sample tray in the MonoTemper Monolith NT.115 instrument. Sixteen PCR tubes were prepared, with 10 μL DMSO added to PCR tubes 2-16 and 20 μL of 5000 μM quinofumelin (or phenamacril) dissolved in DMSO added to PCR tube 1. Then, 10 μL of quinofumelin from PCR tube 1 was transferred and thoroughly mixed with PCR tube 2 using a pipette. These steps were repeated for PCR tubes 3-16. Following mixing with PCR tube 16 (discarding the remaining 10 μL), each PCR tube was thoroughly mixed with 190 μL DHODHII-GFP protein, and incubated at room temperature for 30 min. Monolith NT.115 standard capillaries were inserted into each PCR tube to allow solution entry into the capillary. The capillaries 1-16 were subsequently placed in the sample tray of the device and the program was initiated. All measurements were performed using 80% (Auto-detect) Nano-RED. MonoTemper Monolith NT.115 software was used to evaluate the binding affinity between the target protein and quinofumelin using the Kd values and the signal to noise ratio, and the results were combined and analyzed using Mo. Affinity Analysis v2.3 software (Rotem et al., 2015).

Statistical analysis

The data in the study were analyzed using analysis of variance (ANOVA) with the SPSS 14.0 software (SPSS Inc. Chicago, IL, USA). When ANOVA yielded a significant result (p = 0.05), means were compared using Fisher’s protected least significant difference (PLSD).